Extracting Angular Observables without a Likelihood\\and Applications to Rare Decays

### 2014-07-28 draft.tex
confirming O(1)
Fix some of the issues in App. E found by Fred

In order to study the performance of the proposed method, we simulate measurements for one uni-angular
and one tri-angular decay distribution for various numbers of events. As inputs we use the SM predictions
for the decays $B\to K\ell^+\ell^-$ and $B\to K^*(\to K\pi)\ell^+\ell^-$, respectively. We cover the
$q^2$ ranges $1\GeV^2$ to $6\GeV^2$, and $15\GeV^2$ to to the respective kinematic endpoint, with two
bin widths: $1\GeV^2$ and $0.5\GeV^2$. This is to ensure that a wide spectrum of possible values for
the angular observables. We find:
\begin{itemize}
    \item In all studied cases we observed not a single bias in the pull distribution of the observable.
    \item The error $\sigma_k(N)$ of an angular observable $S_k$, evaluate for $N$ events is very well
        fitted by
        \begin{equation}
            \sigma_k(N) = \frac{\sigma_k(1)}{\sqrt{N}}
        \end{equation}
        with $\sigma_k(1) = \order{1}$. \todo{Check this!}
    \item \todo{Plot Skewness and Kurtosis for $S_5$ and $S_7$ as functions of $N$.}
    \item \todo{Perform Mardia's Skewness and Kurtosis tests.}
    \item \dots
\end{itemize}
We find that all of the above results show consistenly convergence of the distribution for the angular observables
towards a multivariate normal distribution.

@@

We propose the decomposition of $S$ in terms of the associated Legendre polynomials $p_{j}^{(m)}(\cos\theta)$, with total angular momentum $j$ and its third component $m=m_1 + m_2$.
\begin{equation}
    S(\vec{\nu}, \theta) = \sum_j S_{j,m}(\vec{\nu}) p_{j}^{(m)}(\cos\theta)\,.
\end{equation}
This parametrization has two merits. First, we can immediately project out the angular observables $S_{j,m}$ by means of \refeq{legendre-scalar-product}:
\begin{equation}
    S_{j,m}(\vec{\nu}) = \frac{1}{n_{j,m}} \int_{-1}^{+1} \rmdx{\cos\theta} S(\vec{\nu},\cos \theta) p_{j}^{(m)}(\cos\theta)\,.
\end{equation}
Second, we can immediately express $S_{j,m}(\vec{\nu})$ in terms of
the partial-wave amplitudes, \fred{We may exchage the integral and the
  series because each element of the series is a product of
  polynomials on the compact support [-1,1], and thus each integral is
  absolutely convergent}
\begin{equation}
    \label{eq:partial-wave-observable-infinite}
    \begin{aligned}
        S_{j,m}(\vec{\nu})
            & = \frac{1}{n_{j,m}} \int_{-1}^{+1} \rmdx{\cos\theta} p_{j}^{(m_1 + m_2)}(\cos\theta) \sum_{l_1,l_2=0}^\infty F\left[A_{1}^{l_1} p_{l_1}^{(m_1)}(\cos\theta) A_{2}^{*(l_2)}p_{l_2}^{(m_2)}(\cos\theta)\right]\\
            & = \sum_{l_1,l_2=0}^\infty F\left[A_{1}^{l_1} A_{2}^{*(l_2)}\right] \frac{T_{l_1,l_2,j}^{(m_1,m_2)}}{n_{j,m}}\,,\qquad\text{with }m = m_1 + m_2\,.
    \end{aligned}
\end{equation}
In the last step, we use Gaunt's formula \cite{Gaunt:1929} to integrate a triple product of associated Legendre polynomials,
\begin{equation}
\begin{aligned}
    T_{l_1,l_2,j}^{(m_1,m_2)}
        & = \int_{-1}^{+1} \rmdx{\cos\theta} p_{j}^{(m_1 + m_2)}(\cos\theta) p_{l_1}^{(m_1)}(\cos\theta) p_{l_2}^{(m_2)}(\cos\theta)\\
        & = (-1)^{s - l_1 - m_2} \frac{2 (l_1 + m_1)! (l_2 + m_2)! (2s - 2 l_2)! s!}{(l_1 - m_1)! (s - j)! (s - l_1)! (s - l_2)! (2s + 1)!}\\
        & \times \sum_{t=p}^q (-1)^t \frac{(j + m + t)!(l_1 + l_2 - m - t)!}{t! (j - m - t)! (l_1 - l_2 + m + t)! (l_2 - m_2 - t)!}\,,
\end{aligned}
\end{equation}
where
\begin{equation}
\begin{aligned}
    m & = m_1 + m_2\,, &
    m_1 & \geq m_2\,,  \\
    j, l_1, l_2 & \geq 0\,, &
    m, m_1, m_2 & \geq 0\,,
\end{aligned}
\end{equation}
and
\begin{equation}
\begin{aligned}
    s & = \frac{j + l_1 + l_2}{2}\,, &
    p & = \max(0, l_2 - l_1 - m)\,, &
    q & = \min(l_1 + l_2 - m, j - m, l_2 - m_2)\,.
\end{aligned}
\end{equation}
The necessary conditions for $T \neq 0$ are
\begin{equation}
    \label{eq:angular-momentum-addition}
    s \in \mathbb{N}\qquad \wedge \qquad l_1 - l_2 \leq l \leq l_1 + l_2,.
\end{equation}
The latter condition is well known from adding angular momenta. Note, however, that
the sum in \refeq{def-partial-wave-observable} goes to infinitely high angular momenta $l_1$ and $l_2$. As a consequence
of this and of \refeq{angular-momentum-addition}, the angular observables $P_{j,m}$
consist of sums with infinitely many terms. It is then up to theoretical analyses to
estimate or calculate the impact of individual contributions, with or without the cut off $L$.
For such an analysis in the decay $B\to K\pi\ell^+\ell^-$ we refer to \cite{Das:2014sra}.
We close by remarking that for any angular-momentum cut off $\ell < L$\fred{Do you mean $L < \infty$? Don't use $\ell$ again}, one can find an upper limit for the squared sum
of all remaining partial waves through the inclusive decay; i.e., schematically
\begin{equation}
    \sum_{i=1}^2 \sum_{l=L}^\infty |A_i^{(l)}|^2 \leq \frac{\rmdx{^{1 + \dim\vec\nu}} \Gamma_\text{incl}}{\rmdx{\cos\theta} \rmdx{\vec\nu}}\,.
\end{equation}

>>>>

In order to study the performance of the proposed method, we simulate measurements for one uni-angular
and one tri-angular decay distribution for various numbers of events. As inputs we use the SM predictions
for the decays $B\to K\ell^+\ell^-$ and $B\to K^*(\to K\pi)\ell^+\ell^-$, respectively. We cover the
$q^2$ ranges $1\GeV^2$ to $6\GeV^2$, and $15\GeV^2$ to to the respective kinematic endpoint, with two
bin widths: $1\GeV^2$ and $0.5\GeV^2$. This is to ensure that a wide spectrum of possible values for
the angular observables. We find:
\begin{itemize}
    \item In all studied cases we observed not a single bias in the pull distribution of the observable.
    \item The error $\sigma_k(N)$ of an angular observable $S_k$, evaluate for $N$ events is very well
        fitted by
        \begin{equation}
            \sigma_k(N) = \frac{\sigma_k(1)}{\sqrt{N}}
        \end{equation}
        with $\sigma_k(1) = \order{1}$. 
    \item \todo{Plot Skewness and Kurtosis for $S_5$ and $S_7$ as functions of $N$.}
    \item \todo{Perform Mardia's Skewness and Kurtosis tests.}
    \item \dots
\end{itemize}
We find that all of the above results show consistenly convergence of the distribution for the angular observables
towards a multivariate normal distribution.

@@

We propose the decomposition of $S$ in terms of the associated Legendre polynomials $p_{j}^{(m)}(\cos\theta)$, with total angular momentum $j$ and its third component $m=m_1 + m_2$.
\begin{equation}
    S(\vec{\nu}, \theta) = \sum_j S_{j,m}(\vec{\nu}) p_{j}^{(m)}(\cos\theta)\,.
\end{equation}
This parametrization has two merits. First, we can immediately project out the angular observables $S_{j,m}$ by means of \refeq{legendre-scalar-product}:
\begin{equation}
    S_{j,m}(\vec{\nu}) = \frac{1}{n_{j,m}} \int_{-1}^{+1} \rmdx{\cos\theta} S(\vec{\nu},\cos \theta) p_{j}^{(m)}(\cos\theta)\,.
\end{equation}
(Here and in the next step we may exchage the integral and the
series because each element of the series is a product of
polynomials on the compact support [-1,1], and thus each integral
is absolutely convergent).
Second, we can immediately express $S_{j,m}(\vec{\nu})$ in terms of
the partial-wave amplitudes,
\begin{equation}
    \label{eq:partial-wave-observable-infinite}
    \begin{aligned}
        S_{j,m}(\vec{\nu})
            & = \frac{1}{n_{j,m}} \int_{-1}^{+1} \rmdx{\cos\theta} p_{j}^{(m_1 + m_2)}(\cos\theta) \sum_{l_1,l_2=0}^\infty F\left[A_{1}^{l_1} p_{l_1}^{(m_1)}(\cos\theta) A_{2}^{*(l_2)}p_{l_2}^{(m_2)}(\cos\theta)\right]\\
            & = \sum_{l_1,l_2=0}^\infty F\left[A_{1}^{l_1} A_{2}^{*(l_2)}\right] \frac{T_{l_1,l_2,j}^{(m_1,m_2)}}{n_{j,m}}\,,\qquad\text{with }m = m_1 + m_2\,.
    \end{aligned}
\end{equation}

In the last step, we use Gaunt's formula \cite{Gaunt:1929} to integrate a triple product of associated Legendre polynomials,
\begin{equation}
\begin{aligned}
    T_{l_1,l_2,j}^{(m_1,m_2)}
        & = \int_{-1}^{+1} \rmdx{\cos\theta} p_{j}^{(m_1 + m_2)}(\cos\theta) p_{l_1}^{(m_1)}(\cos\theta) p_{l_2}^{(m_2)}(\cos\theta)\\
        & = (-1)^{s - l_1 - m_2} \frac{2 (l_1 + m_1)! (l_2 + m_2)! (2s - 2 l_2)! s!}{(l_1 - m_1)! (s - j)! (s - l_1)! (s - l_2)! (2s + 1)!}\\
        & \times \sum_{t=p}^q (-1)^t \frac{(j + m + t)!(l_1 + l_2 - m - t)!}{t! (j - m - t)! (l_1 - l_2 + m + t)! (l_2 - m_2 - t)!}\,,
\end{aligned}
\end{equation}
where
\begin{equation}
\begin{aligned}
    m & = m_1 + m_2\,, &
    m_1 & \geq m_2\,,  \\
    j, l_1, l_2 & \geq 0\,, &
    m, m_1, m_2 & \geq 0\,,
\end{aligned}
\end{equation}
and
\begin{equation}
\begin{aligned}
    s & = \frac{j + l_1 + l_2}{2}\,, &
    p & = \max(0, l_2 - l_1 - m)\,, &
    q & = \min(l_1 + l_2 - m, j - m, l_2 - m_2)\,.
\end{aligned}
\end{equation}
The necessary conditions for $T \neq 0$ are
\begin{equation}
    \label{eq:angular-momentum-addition}
    s \in \mathbb{N}\qquad \wedge \qquad l_1 - l_2 \leq l \leq l_1 + l_2,.
\end{equation}
The latter condition is well known from adding angular momenta. Note, however, that
the sum in \refeq{def-partial-wave-observable} goes to infinitely high angular momenta $l_1$ and $l_2$. As a consequence
of this and of \refeq{angular-momentum-addition}, the angular observables $P_{j,m}$
consist of sums with infinitely many terms. It is then up to theoretical analyses to
estimate or calculate the impact of individual contributions, with or without the cut off $L$.
For such an analysis in the decay $B\to K\pi\ell^+\ell^-$ we refer to \cite{Das:2014sra}.
We close by remarking that for any angular-momentum cut off $l < L$\fred{Do you mean $L < \infty$?}, one can find an upper limit for the squared sum
of all remaining partial waves through the inclusive decay; i.e., schematically
\begin{equation}
    \sum_{i=1}^2 \sum_{l=L}^\infty |A_i^{(l)}|^2 \leq \frac{\rmdx{^{1 + \dim\vec\nu}} \Gamma_\text{incl}}{\rmdx{\cos\theta} \rmdx{\vec\nu}}\,.
\end{equation}
